---
title: "Lab 6: Nothing is Normal About This!"
author: Samantha Cavalli and Bri Cirillo
date: "9 April 2020"
output: html_notebook
---

```{r}
#install.packages("Deriv")
library(Deriv)
```

https://github.com/s1173652/Lab6

# Part I: The Rightful pgauss()

## Problem 1

```{r}
#' Compute F(x) for a Gaussian random variable 
#' @param x a vector of values to be evaluated at
#' @param mean the mean of the random variable
#' @param var the variance of the random variable 
#'
#' @return the cumulative distribution function evaluated at all the values of x

x<- c(1,2,3)

pgauss <- function(x,mean=0, var=1){
  
  cum.vector <- vector(length = length(x))
  
  for (i in 1:length(x)) {
    
  cum.vector[i] <- integrate(f=function(x) dnorm(x,0,1), lower= -Inf, upper= x[i])[[1]]
    
  
  }
  
  return(cum.vector)
}

pgauss(x,mean=0,var=1) 

#testing the function to understand how it works
```

## Problem 2

```{r}
pgauss(3, mean=1, var=2)

#finding the probability that a Gaussian RV with mean µ = 1 and variance = 2 exceeds 3
```

The probability that a Gaussian random variable with mean = 1 and variance = 2 exceeds the value of 3 is 1 - 0.999, which is 0.001.

```{r}
pnorm(q=3,mean=0,sd=1)

# using pnorm to check our prior result using pgauss
```

We can verify our result using pnorm and get the same answer. 

# Part II: What's a Normal Score?

## Problem 3

```{r}
curve(pnorm(x,5,10), from = -15, to= 25)

# plotting a cdf curve for a Gaussian RV with mean=5 and var=100
```

It looks like the 0.25 quantile of this distribution is at is about -2. 

## Problem 4

```{r}
#' Compute the quantile function for a Gaussian random variable 
#' @param q a vector of values to be evaluated at
#' @param mean the mean of the random variable
#' @param var the variance of the random variable 
#'
#' @return the cumulative distribution function evaluated at all the values of x

qgauss <- function(q, mean=0, var=1){
  
  quant.vector <- vector(length = length(q))
  
  for (i in 1: length(q)){
    f <- function(x) pgauss(x, mean,var) - q[i]
     quant.vector[i] <- uniroot(f, c(mean-3*sqrt(var), mean+3*sqrt(var)))[[1]]
    
  }
    
 return(quant.vector) 
}

qgauss(.25,mean=0,var=1)

#testing qguass to better understand the function
```

## Problem 5

```{r}
qgauss(0.25, mean = 5, var = 100)

# finding the 0.25 quantile of a GRV with mean=5 and var=100

qnorm(0.25,mean=5,sd=10)

# checking the previous answer with qnorm
```

The 0.25 quantile is -1.745. This answer matches the graphical solution found in Problem 3. 

# Part III: What's the Most Likely Mean?

## Problem 6

```{r}
#' Compute the log-likelihood function for a Gaussian random variable 
#' @param mean the mean of the random variable
#' @param var the variance of the random variable
#' @param data a vector of values to be evaluated at
#'
#' @return the cumulative distribution function evaluated at all the values of x

data <- c(1,2,3)

gaussian.ll <- function(mean,var,data){
  
     n <- length(data)
    
    sum.data <- sum((data-mean)^2)
    
    ll <- -(n/2)*log(2*pi*var)-(1/(2*var))*(sum.data)
    
  return(ll)
}

gaussian.ll(data=c(1,2,3), mean=0, var=1)

#testing guassian.11 to better understand the function
```

## Problem 7

```{r}
gaussian.ll.v <- Vectorize(gaussian.ll, vectorize.args = 'mean')

# vectorizing the mean argument
```

The mean argument is vectorized.

## Problem 8

```{r}
set.seed(1)

x <- rnorm(100, mean = 0, sd = 1)

# creating a random sample of 100 from a Gaussian distribution

curve(gaussian.ll.v(mean,1,x), from = -5, to = 5, xname= 'mean')
abline(v=0.1088874, lty =3)

# plot of the log-likelihood function
```

The value of the mean that maximizes the log-likelihood is around 0. This makes sense given that we generated the sample with a normal distribution, where the standard mean is 0 and the MLE of a normal random variable is the sample mean.

## Problem 9

```{r}
optimize(gaussian.ll.v,var=1, data= x, interval=c(-5,5),maximum = TRUE)

# finding the value of the mean that maximizes the log-liklihood function
```

The maximum likelihood estimator for the function is 0.1089. 

## Problem 10

```{r}
mean(x)

# computing the mean of x
```

The mean of x is equal to 0.1089, which is the same maximizer found in problem 9. This result is what we would expect since the MLE of a normal random sample is the sample mean. 


